---
title: "Analysis"
output:
  pdf_document:
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Behavioural Research Statistical Analysis

## Testing Event Boundary Memory Hypotheses

This notebook analyses the cleaned recognition dataset to test three hypotheses regarding memory performance under **Natural Cut (NB)** and **Abrupt Cut (AB)** conditions.

Additionally, two cognitive memory metrics are computed:

* **Recognition Memory Index (REC)**
* **Lure Discrimination Index (LDI)**

---

# Load Libraries

```{r}
library(tidyverse)
library(stringr)
```

---

# Load Clean Dataset

```{r}
data <- read.csv("final_cleaned_recognition_data.csv", check.names = FALSE)

# Convert variables

data$Condition <- factor(data$`AB/NB`)
data$Frames <- factor(data$Frames)
data$Subject_id <- factor(data$Subject_id)

# Accuracy and RT

data$accuracy <- as.numeric(data$resp.corr)
data$rt <- as.numeric(data$resp.rt)
```

---

# Hypothesis 1

## NB participants should show higher accuracy and faster RT

## Participant-level summary

```{r}
summary_data <- data %>%
  group_by(Subject_id, Condition) %>%
  summarise(
    mean_accuracy = mean(accuracy, na.rm = TRUE),
    mean_rt = mean(rt, na.rm = TRUE),
    .groups = "drop"
  )
```

---

## Statistical Tests

```{r}
# Accuracy comparison
t.test(mean_accuracy ~ Condition, data = summary_data)

# Reaction time comparison
t.test(mean_rt ~ Condition, data = summary_data)
```

---

## Visualization

### Recognition Accuracy by Condition

```{r}
ggplot(summary_data, aes(x = Condition, y = mean_accuracy, fill = Condition)) +
  stat_summary(fun = mean, geom = "bar") +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.2) +
  labs(title = "Recognition Accuracy by Condition",
       x = "Condition",
       y = "Mean Accuracy") +
  theme_minimal()
```

### Reaction Time Distribution

```{r}
ggplot(summary_data, aes(x = Condition, y = mean_rt, fill = Condition)) +
  geom_violin(trim = FALSE, alpha = 0.7) +
  geom_boxplot(width = 0.1, fill = "white") +
  labs(title = "Reaction Time Distribution",
       x = "Condition",
       y = "Reaction Time (seconds)") +
  theme_minimal()
```

---

# Hypothesis 2

## BB frames should show higher accuracy in NB compared to AB

```{r}
frame_summary <- data %>%
  group_by(Subject_id, Condition, Frames) %>%
  summarise(
    mean_accuracy = mean(accuracy, na.rm = TRUE),
    .groups = "drop"
  )
```

---

## Two-way ANOVA

```{r}
anova_result <- aov(mean_accuracy ~ Condition * Frames, data = frame_summary)
summary(anova_result)
```

---

## Visualization

```{r}
ggplot(frame_summary, aes(x = Frames, y = mean_accuracy, fill = Condition)) +
  stat_summary(fun = mean, geom = "bar", position = "dodge") +
  stat_summary(fun.data = mean_se,
               geom = "errorbar",
               position = position_dodge(0.9),
               width = 0.2) +
  labs(title = "Accuracy by Frame Type and Condition",
       x = "Frame Type",
       y = "Mean Accuracy") +
  theme_minimal()
```

---

# Hypothesis 3

## EM frames should not differ significantly between conditions

```{r}
em_data <- frame_summary %>%
  filter(Frames == "EM")

# t-test
t.test(mean_accuracy ~ Condition, data = em_data)
```

---

# Confidence vs Correctness

```{r}
ggplot(data, aes(x = factor(accuracy), y = confidence)) +
  geom_violin(trim = FALSE, fill = "skyblue") +
  geom_boxplot(width = 0.1, fill = "white") +
  labs(title = "Confidence vs Correctness",
       x = "Correct (1) vs Incorrect (0)",
       y = "Confidence") +
  theme_minimal()
```

---

# Recognition Memory Metrics

## Recognition Memory Index (REC)

Formula:

REC = P("old" | Target) − P("old" | Foil)

```{r}
# Probability participant correctly identified target
p_old_target <- mean(data$resp.corr == 1, na.rm = TRUE)

# Probability participant incorrectly identifies foil
p_old_foil <- mean(data$resp.corr == 0, na.rm = TRUE)

REC <- p_old_target - p_old_foil

REC
```

---

# Lure Discrimination Index (LDI)

Formula:

LDI = P("similar" | Lure) − P("similar" | Foil)

```{r}
# Example approximation using frame types
p_similar_lure <- mean(data$Frames == "BB" & data$accuracy == 0, na.rm = TRUE)
p_similar_foil <- mean(data$Frames == "EM" & data$accuracy == 0, na.rm = TRUE)

LDI <- p_similar_lure - p_similar_foil

LDI
```

---

# Summary

This analysis tested three hypotheses:

1. **Natural cut videos improve recognition accuracy and reaction time.**
2. **Before-boundary frames show stronger memory in the natural condition.**
3. **Event-middle frames show little difference between conditions.**

Additionally, the notebook computes **Recognition Memory Index (REC)** and **Lure Discrimination Index (LDI)** to quantify memory discrimination ability.


# Experiment Analysis

```{r}
library(tidyverse)

# -----------------------------------------
# Number of participants in each condition
# -----------------------------------------

participant_count <- data %>%
  distinct(Subject_id, Condition) %>%
  count(Condition)

participant_count


# -----------------------------------------
# Frames per participant (BB vs EM)
# -----------------------------------------

frame_counts <- data %>%
  group_by(Subject_id, Frames) %>%
  summarise(
    n_frames = n(),
    .groups = "drop"
  ) %>%
  pivot_wider(
    names_from = Frames,
    values_from = n_frames,
    values_fill = 0
  )

frame_counts


# -----------------------------------------
# Optional: total BB and EM counts overall
# -----------------------------------------

total_frames <- data %>%
  count(Frames)

total_frames
```

```{r}
library(tidyverse)

# Ensure numeric types
data$confidence <- as.numeric(data$confidence)
data$accuracy <- as.numeric(data$resp.corr)
data$rt <- as.numeric(data$resp.rt)

# ------------------------------------------
# Summary statistics by confidence level
# ------------------------------------------

confidence_summary <- data %>%
  group_by(confidence) %>%
  summarise(
    n_trials = n(),
    accuracy_rate = mean(accuracy, na.rm = TRUE),
    avg_response_time = mean(rt, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(confidence)

print(confidence_summary)

# ------------------------------------------
# Visualization 1: Accuracy vs Confidence
# ------------------------------------------

ggplot(confidence_summary, aes(x = confidence, y = accuracy_rate)) +
  geom_line(size = 1) +
  geom_point(size = 3) +
  labs(
    title = "Recognition Accuracy by Confidence Level",
    x = "Confidence Rating",
    y = "Accuracy Rate"
  ) +
  theme_minimal()

# ------------------------------------------
# Visualization 2: Response Time vs Confidence
# ------------------------------------------

ggplot(confidence_summary, aes(x = confidence, y = avg_response_time)) +
  geom_line(size = 1) +
  geom_point(size = 3) +
  labs(
    title = "Average Response Time by Confidence Level",
    x = "Confidence Rating",
    y = "Average Response Time (seconds)"
  ) +
  theme_minimal()
```

```{r}
library(tidyverse)

# Ensure proper formats
data$resp.keys <- as.character(data$resp.keys)
data$accuracy <- as.numeric(data$resp.corr)
data$rt <- as.numeric(data$resp.rt)
data$confidence <- as.numeric(data$confidence)

# ------------------------------------------
# Summary by response key
# ------------------------------------------

key_summary <- data %>%
  group_by(resp.keys) %>%
  summarise(
    n_trials = n(),
    accuracy_rate = mean(accuracy, na.rm = TRUE),
    avg_response_time = mean(rt, na.rm = TRUE),
    .groups = "drop"
  )

print(key_summary)

# ------------------------------------------
# Most selected confidence level for r and l
# ------------------------------------------

confidence_mode <- data %>%
  group_by(resp.keys, confidence) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(resp.keys) %>%
  slice_max(n, n = 1)

print(confidence_mode)

# ------------------------------------------
# Visualization 1: Accuracy by response key
# ------------------------------------------

ggplot(key_summary, aes(x = resp.keys, y = accuracy_rate, fill = resp.keys)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Accuracy by Response Key",
    x = "Response Key",
    y = "Accuracy Rate"
  ) +
  theme_minimal()

# ------------------------------------------
# Visualization 2: Response time by key
# ------------------------------------------

ggplot(key_summary, aes(x = resp.keys, y = avg_response_time, fill = resp.keys)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Average Response Time by Response Key",
    x = "Response Key",
    y = "Average Response Time (seconds)"
  ) +
  theme_minimal()

# ------------------------------------------
# Visualization 3: Confidence distribution
# ------------------------------------------

ggplot(data, aes(x = factor(confidence), fill = resp.keys)) +
  geom_bar(position = "dodge") +
  labs(
    title = "Confidence Level Distribution by Response Key",
    x = "Confidence Level",
    y = "Count"
  ) +
  theme_minimal()
```
