---
title: "SRS"
output: pdf_document
date: "2026-03-10"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Event Boundary Disruption and Recognition Memory

This analysis investigates whether disrupting event boundaries in videos affects recognition memory performance.

Three hypotheses are tested:

1. Participants who watched **Naturally cut videos (NB)** will show higher recognition accuracy and faster response times than those who watched **Abruptly cut videos (AB)**.

2. **Before-boundary (BB) frames** will show higher recognition accuracy in the Natural Cut condition compared to the Abrupt Cut condition.

3. Recognition performance for **Event-middle (EM) frames** will not differ significantly between the two conditions.

---

## Load Required Libraries

```{r}
library(tidyverse)
library(ggpubr)
library(rstatix)
```

---


```{r}
library(tidyverse)
library(stringr)

# Load dataset
data <- read.csv("/Users/rohit.rohon01gmail.com/Downloads/cleaned_recognition_data.csv",
                 check.names = FALSE,
                 stringsAsFactors = FALSE)

# Clean whitespace across all character columns
data <- data %>%
  mutate(across(where(is.character), trimws))

# Create Condition column
data$Condition <- factor(data$`NB.AB`)

# Convert important columns to numeric
data$accuracy <- as.numeric(data$resp.corr)
data$rt <- as.numeric(data$resp.rt)

# Extract Frame Type (BB or EM) from filename
data$FrameType <- ifelse(grepl("_BB_", data$target_img), "BB", "EM")
data$FrameType <- factor(data$FrameType)

# Convert participant to factor
data$participant <- factor(data$participant)

# Remove rows with missing accuracy
data <- data %>% filter(!is.na(accuracy))

# Verify Condition groups
table(data$Condition)

```


```{r}
# ---- Hypothesis 1 ----

library(tidyverse)

summary_data <- data %>%
  group_by(participant, Condition) %>%
  summarise(
    mean_accuracy = mean(accuracy, na.rm = TRUE),
    mean_rt = mean(rt, na.rm = TRUE),
    .groups = "drop"
  )

# Statistical tests
t.test(mean_accuracy ~ Condition, data = summary_data)
t.test(mean_rt ~ Condition, data = summary_data)

# Accuracy by condition (bar chart)
ggplot(summary_data, aes(x = Condition, y = mean_accuracy, fill = Condition)) +
  stat_summary(fun = mean, geom = "bar") +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.2) +
  labs(
    title = "Recognition Accuracy by Condition",
    x = "Condition",
    y = "Mean Accuracy"
  ) +
  theme_minimal()

# Reaction time distribution (violin plot)
ggplot(summary_data, aes(x = Condition, y = mean_rt, fill = Condition)) +
  geom_violin(trim = FALSE, alpha = 0.7) +
  geom_boxplot(width = 0.1, fill = "white") +
  labs(
    title = "Reaction Time Distribution by Condition",
    x = "Condition",
    y = "Reaction Time (seconds)"
  ) +
  theme_minimal()
```

```{r}
# ---- Hypothesis 2 ----

frame_summary <- data %>%
  group_by(participant, Condition, FrameType) %>%
  summarise(
    mean_accuracy = mean(accuracy, na.rm = TRUE),
    .groups = "drop"
  )

# Two-way ANOVA
anova_result <- aov(mean_accuracy ~ Condition * FrameType, data = frame_summary)
summary(anova_result)

# Grouped bar chart for interaction
ggplot(frame_summary, aes(x = FrameType, y = mean_accuracy, fill = Condition)) +
  stat_summary(fun = mean, geom = "bar", position = "dodge") +
  stat_summary(fun.data = mean_se, geom = "errorbar",
               position = position_dodge(0.9), width = 0.2) +
  labs(
    title = "Accuracy by Frame Type and Condition",
    x = "Frame Type",
    y = "Mean Accuracy"
  ) +
  theme_minimal()
```


```{r}
# ---- Hypothesis 3 ----

em_data <- frame_summary %>%
  filter(FrameType == "EM")

# t-test
t.test(mean_accuracy ~ Condition, data = em_data)

# Accuracy comparison for EM frames (bar plot)
ggplot(em_data, aes(x = Condition, y = mean_accuracy, fill = Condition)) +
  stat_summary(fun = mean, geom = "bar") +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.2) +
  labs(
    title = "Recognition Accuracy for Event-Middle Frames",
    x = "Condition",
    y = "Mean Accuracy"
  ) +
  theme_minimal()

# Confidence vs correctness (violin plot)
ggplot(data, aes(x = factor(accuracy), y = confidence)) +
  geom_violin(trim = FALSE, fill = "skyblue", alpha = 0.7) +
  geom_boxplot(width = 0.1, fill = "white") +
  labs(
    title = "Confidence vs Correctness",
    x = "Correct (1) vs Incorrect (0)",
    y = "Confidence Rating"
  ) +
  theme_minimal()
```